---
title: "How to download high resolution satellite images with R"
author: "Edson Silva-Júnior"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# How to download high resolution satellite images with R

Our first step is to library the required packages: 

- We use [geobr](https://github.com/ipeaGIT/geobr) package to download our area shapefile, Recife municipality, Pernambuco's capital; 

- We use [tidyverse](https://www.tidyverse.org/packages) to trate and filter our shapefile, to get only Recife area feature;

- We use [maptiles](https://github.com/riatelab/maptilesmaptiles Github page) package to download our satellite image, onde raster class;

- We use [tidyterra](https://dieghernan.github.io/tidyterra) package to visualizing multicolar raster throught ggplot;

- We use [terra](https://github.com/ipeaGIT/geobr) package to export our downloaded raster.

# Packages 

```{r}
library(geobr)

library(tidyverse)

library(maptiles)

library(tidyterra)

library(terra)
```

# Base Shapefile

## Importing 

First, lets to download a brazilian municipalities shapefile throught `geobr::read_municipality()` and use `dplyr::filter()` to get only Recife feature.

```{r}
recife <- geobr::read_municipality(year = 2019) |> 
  dplyr::filter(name_muni == "Recife")

```

## Visualizing 

Next, lets vicualize our base shapefile, using a ggplot plot.

```{r}
recife

ggplot() +
  geom_sf(data = recife, color = "black", linewidth = 1)
```

# Satellite image

## Downloading

Now, our next step is to download our satellite raster, to Recife's area. We use `maptiles::get_tiles()`, informing:

- Our shapefile to use as base area (`x`);

- Which imagery we want to use (`provider`). `"Esri.WorldImagery"` let us to use a raster such as Google Earth. Check [maptiles Github page](https://github.com/riatelab/maptilesmaptiles Github page) to see more imageries options;

- Zoom (`zoom`), to set how resoluted the raster let be. See [OpenStreetMap wiki](https://wiki.openstreetmap.org/wiki/Zoom_levels) for details.

```{r}
image <- maptiles::get_tiles(x = recife,
                              provider = "Esri.WorldImagery",
                              zoom = 16)
```


## Visualizing 

Finaly, we obtained our satellite raster, using `tidyterra::geom_spatraster_rgb()` to load and visualize the colored raster. Check [tidyterra](https://dieghernan.github.io/tidyterra) for details.

```{r}
image

ggplot() +
  tidyterra::geom_spatraster_rgb(data = image) +
  geom_sf(data = recife, fill = "transparent", color = "red", linewidth = 1) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme_bw()
```

## Exporting

To the end, we go to export our raster, mainly to do not be necessary download our satellite image again, using `terra::writeRaster()`

```{r}
image |>
  terra::writeRaster("recife_image.tif")
```

